Challenges in Developing Great Quasi-Monte Carlo Software

### 2023-03-17 MCQMC2022ChallengesQMCSoftwareArticle.tex
Updates from Overleaf
Fred 2023 Mac 14 Updates
aleksei's small changes and sign off

%%%%%%% Author sign off on this first submission, last revision
% Sou-Cheng Choi signed off on
% Yuhan Ding signed off on
% Fred J. Hickernell signed off on
% Jagadeeswaran Rathinavel signed off on
% Aleksei Sorokin signed off on

@@

\abstract{
Quasi-Monte Carlo (QMC) methods have developed over several decades. With the explosion in computational science, there is a need for great software that implements QMC algorithms. We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.  We illustrate these criteria and steps in practice with Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
}

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu_n} := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with a small number of sampling points, $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu_n}$:
\begin{gather}
\label{eq:cubSummary}
	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=1}^{n} f(\bsx_i) =: \hat{\mu}_n, \\
 \label{CDHJS_eq:error_tol}
 \abs{\mu - \hat{\mu}_n} \le \varepsilon.
\end{gather}
Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.  Great QMC software environments allow different implementations of these components to be freely interchanged.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \url{https://github.com/QMCSoftware/QMCSoftware/blob/develop/demos/talk_paper_demos/MCQMC2022_Article_Figures.ipynb}  }
    \label{fig:beam}
\end{figure}

>>>>

%%%%%%% Author sign off on this first submission, last revision
% Sou-Cheng Choi signed off on
% Yuhan Ding signed off on
% Fred J. Hickernell signed off on
% Jagadeeswaran Rathinavel signed off on
% Aleksei Sorokin signed off on 3/17/2023

@@

\abstract{
Quasi-Monte Carlo (QMC) methods have developed over several decades. With the explosion in computational science, there is a need for great software that implements QMC algorithms. We summarize the  QMC software that has been developed to date, propose some criteria for developing great QMC software, and suggest some steps toward achieving great software.  We illustrate these criteria and steps in practice with the Quasi-Monte Carlo Python library (QMCPy), an open-source community software framework, extensible by design with common programming interfaces to an increasing number of existing or emerging QMC libraries developed by the greater community of QMC researchers.
}

@@

A QMC approximation of $\mu := \EE[f(\bsX)]$, $\bsX \sim \calU[0,1]^d$ can be implemented in a few steps:
\begin{enumerate}
    \item Draw a sequence of $n$ low-discrepancy (LD) \cite{Nie92, SloJoe94, DicEtal14a, DicEtal22a} nodes,
    $\bsx_1,\dots,\bsx_n$ that mimic $\calU[0,1]^d$.
    \item Evaluate the integrand $f$ at these nodes to obtain $f(\bsx_i)$, $i=1,\dots,n$.
    \item Estimate the true mean, $\mu$, by the sample mean,
    \begin{equation}
        \hat{\mu}_n := \frac{1}{n} \sum_{i=1}^n f(\bsx_i).
        \label{eq:mc_average}
    \end{equation}
\end{enumerate}
However, the practice of QMC is often more complicated.  The original problem may need to be rewritten to fit the above form and/or to facilitate a good approximation with a small number of sampling points, $n$.  Practitioners may wish to determine $n$ adaptively to satisfy a prescribed error tolerance.

@@

Expanding on the problem formulation in the introduction, we want to approximate well the integral or expectation, $\mu$, by the sample mean, $\hat{\mu}_n$:
\begin{gather}
\label{eq:cubSummary}
	\mu : = \int_\calT g(\bst) \, \lambda(\bst) \, \D \bst  = \bbE[f(\bsX)] = \int_{[0,1]^d} f(\bsx)  \, \D \bsx \approx \frac 1n \sum_{i=1}^{n} f(\bsx_i) =: \hat{\mu}_n, \\
 \label{CDHJS_eq:error_tol}
 \abs{\mu - \hat{\mu}_n} \le \varepsilon.
\end{gather}
Progressing from the original problem to a satisfactory solution requires several software components, which we describe in the subsections below.  Great QMC software environments allow different implementations of these components to be freely interchanged.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{cantileveredbeamwords.eps}
    \includegraphics[width=1\textwidth]{ldparallelbeam.eps}
    \caption{The expected deflection of a cantilevered beam with random Young's modulus.  QMCPy computes the answer in serial and in parallel using UM-Bridge and docker containers.  For small error tolerances, the time and sample size is $\calO(\varepsilon^{-1})$, and parallel processing gives a speed-up factor of four to five. These plots may be reproduced using the Jupyter notebook \cite{MCQMC2022Figs}.  }
    \label{fig:beam}
\end{figure}
